The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2023-11-29 src/paper/theory.tex
Rewrite conclusion

This paper proposed a new framework to think about the profit allocation in situations concerning a few large and a continuum of smaller players.
The results are especially relevant for upstream-downstream situations, platform settings, and two-sided markets.
The outcomes of the Shapley-value-based gain sharing rule generally concur with intuitive expectations, and the resulting formulas are more tractable than most explicit non-cooperative approaches.

% The illustrative model also highlights the strengths of this approach.
% Despite its simplicity, it highlights that the effects of vertical integration strongly depends on the substitutability of the downstream firms.
% In a time when upstream firms and platforms are increasingly seeking to sell directly to the consumers\footnote{
%     As some recent examples, consider the recent acquisition of Whole Foods by Amazon, Microsoft's ongoing efforts to buy Activision/Blizzard (a major video game publisher), or Tesla's or BMW's plans to sell cars directly without the help of dealerships.
% }
% , such results can be important for competition authorities.

There is a variety of ways in which this paper could be expanded or built upon.
On the theoretical side, a number of the assumptions, such as fringe agents achieving a zero payoff without the platform, could be relaxed,
Besides this, the convergence result for the weighted values could be strengthened if one could manage to show that any sequence along ever more refined partitions of the atomic players converges to the same limit \parencite[à la][]{fogelman1980asymptotic}.

A slightly different direction would be to consider other solution concepts from cooperative game theory in this specific setting.
One obvious candidate is the nucleolus, which shares many of the useful properties of the Shapley-value (it always exists and is unique).
Furthermore, it also has connections to bargaining theory due to it being contained in the core when the latter exists.

The possibilities for practical applications are no less interesting.
The proposed profit sharing framework could be embedded in a more applied model of some real-life setting.
For a step in this direction, see \textcite{stancsics2023hybrid}.

Finally, with time, the cooperative approach to profit distribution could become an almost drop-in alternative of the more usual, but in some sense more extreme assumptions about the bargaining process widely used in industrial organization models.

>>>>

In this paper, I examined the problem of bargaining between one indispensable player and a continuum of fringe players.
I did so using the concept of random order values from cooperative game theory.
I showed that random-order-value-based bargaining has some desirable properties, and that it coincides with intuitive expectations in terms of certain comparative statics.
Furthermore, the resulting distributions of the value are also analytically tractable, and has neat geometric interpretations.

I use random order values to provide a comprehensive framework for a number of important special cases.
Some of those, such as the Shapley-value, or the weighted value in the one-sided case, is already known in the literature.\footnote{
    Aside from looking at it in an abstract manner, my contribution to those cases mostly lie in showing how the fringe's value can be expressed as a kind of expected marginal contribution.
    This is especially relevant in the multi-sided case, where it is important to know how the value not obtained by the indispensible player is distributed among the various sides.
}
Others, such as the weighted value in the multi-sided case, or the interpretation of substitutable "indispensible" players, are, to the best of my knowledge, new.
As I demonstrate, they retain the simplicity and interpretability of the simpler cases, while making the model more general.
Finally, I demonstrate how the proposed framework can be applied in practice by
examining the two-sided platform model of \textcite{armstrong2006competition} with a small extension and a bargaining process over entry fees.
I show that the resulting entry fees have a simple structure, but they also differ from the welfare-maximizing and unilateral price-setting cases in number of interesting ways.
Therefore, in situations where bargaining between players is plausible, certain effects might be missed if one assumes unilateral price-setting.

There is a variety of ways in which this paper could be expanded or built upon.
One obvious direction is to consider other solution concepts from cooperative game theory, such as the nuclelous\footnote{
    It retains many of the desirable properties of the Shapley-value, such as existence and uniqueness, but it is contained in the core when the latter exists, and thus has a more immediate connection to non-cooperative bargaining theory.
}.
Another way to go is to build non-cooperative microfoundations for the random-order-value-based bargaining outcomes.
Many such results exist for the Shapley value, and some also exist for the weighted value, but it might be interesting to see if some of the more general random order values can be motivated in a similar manner.

Besides those theoretical directions, the practical applications are no less interesting.
The proposed profit sharing framework could be embedded in more realistic models of various settings.
The example application in this paper is a small example of that, while \textcite{stancsics2023hybrid} provides a more detailed model involving hybrid platforms.
Due to the simplicity of the results, this approach might be a viable replacement of unilateral price setting in cases where one does not wish to ascribe all the bargaining power to one player.
